---
title: "Remove duplicate using {fastlink}"
output:
  unhcrdown::html_page:
    toc: true
    toc_depth: 2
    toc_float: true
---


```{r echo=TRUE, message=FALSE, warning=FALSE}
# Clear the workspace by removing all objects
rm(list = ls())

# Load the pacman package if it is not already installed.
# if (!requireNamespace("pacman", quietly = TRUE)) {
#   install.packages("pacman")
# }
# 
# pacman::p_load(
#   dplyr,         # Data manipulation and transformation
#   data.table,    # Fast data manipulation with data tables
#   stringi,       # Character string processing
#   lubridate,     # Date and time handling
#   janitor,       # Data cleaning and tabulation functions
#   parallel,      # Parallel computing
#   fastLink,      # Record linkage and deduplication
#   stringdist     # String distance computation
# )

# Get the number of detected cores minus 1
# Reserve one core for non-computational tasks to help prevent system slowdowns or unresponsiveness
numCores <- parallel::detectCores() - 1
```

# Example data frames dfA and  

```{r echo=TRUE, message=FALSE, warning=FALSE}
#data(samplematch)

dfA <- data.frame(
  FIRST_NAME = c("John", "Mary", "Robert", "Michael", "Jennifer", "David", "Karen", "Maria", "Carlos", "James"),
  LAST_NAME = c("Smith III", "Johnson!", "Williams123", "Brown", "Jones", "Davis", "Miller", "Garcia", "Martinez", "Andrson"),
  BIRTH_DATE = c("1981-05-20", "1990-05-15", "1978-12-10", "1985-08-02", "1993-11-25", "1977-03-30", "1988-06-18", "1991-02-05", "1980-09-12", "1982-07-09"),
  gender = c("M", "F", "M", "M", "F", "M", "F", "F", "M", "M")
)

dfB <- data.frame(
  RecipientNameFirst = c("John", "Mary", "Robert", "Michael", "Jennifer", "David", "Karenn", "Carloas", "Mariaa", "James"),
  RecipientNameLast = c("Smith iv", "Brown-", "Williams", "Jones", "John son", "No Name", "Miller", "Martinez", "Garcia", "Anderson"),
  RecipientDateOfBirth = c("1981-05-21", "1992-09-25", "1978-10-12", "1985-08-02", "1993-11-25", "1977-03-30", "1988-06-18", "1980-09-12", "1991-02-05", "1982-07-09"),
  gender = c("M", "F", "M", "M", "F", "M", "F", "M", "F", "M")
)

# # Example data frames dfA and dfB
# dfA <- fread(file = 'dfA.csv',
#              sep = ",",
#              header = T,
#              nThread = numCores)
# 
# dfB <- fread(file = 'dfB.csv',
#              sep = ",",
#              header = T,
#              nThread = numCores)

knitr::kable(dfA)

knitr::kable(dfB)

```


# Assign a unique key ID to each row

```{r echo=TRUE, message=FALSE, warning=FALSE}
dfA <- dfA |>
 dplyr::mutate(row_idA = paste("dfA_", dplyr::row_number(), sep = ""))

dfB <- dfB |>
 dplyr::mutate(row_idB = paste("dfB_", dplyr::row_number(), sep = ""))
```


# Convert the date column to a valid date format.
```{r echo=TRUE, message=FALSE, warning=FALSE}
dfA$BIRTH_DATE <- as.Date(dfA$BIRTH_DATE, format = "%Y-%m-%d")  # "%m/%d/%Y"
dfB$RecipientDateOfBirth <- as.Date(dfB$RecipientDateOfBirth, format = "%Y-%m-%d")  # "%m/%d/%Y"
```


 

# Create new date variables by splitting date of birth into three different parts. 
```{r echo=TRUE, message=FALSE, warning=FALSE}
dfA <- dfA |> 
       dplyr::mutate(dob_day = as.numeric(lubridate::day(BIRTH_DATE)),
                      dob_month = as.numeric(lubridate::month(BIRTH_DATE)),
                      dob_year = as.numeric(lubridate::year(BIRTH_DATE)),
                      DOB=BIRTH_DATE)

dfB <- dfB |> 
       dplyr::mutate(dob_day = as.numeric(lubridate::day(RecipientDateOfBirth)),
                      dob_month = as.numeric(lubridate::month(RecipientDateOfBirth)),
                      dob_year = as.numeric(lubridate::year(RecipientDateOfBirth)),
                      DOB=RecipientDateOfBirth)
```


# Define a function for data cleaning with additional name removal logic
```{r echo=TRUE, message=FALSE, warning=FALSE}
clean_names <- function(names_column) {
  # Step 0: Convert to uppercase
  names_column_new <- toupper(names_column)
  # Step 1: Remove specified name suffixes
  toRemove <- c(" JR", " SR", " IV", " III", " II")
  for (tR in toRemove) {
    names_column_new <- gsub(tR, "", names_column_new)
  }
  # Step 2: Convert special characters to ASCII equivalents
  names_column_new <- iconv(names_column_new, "latin1", "ASCII//TRANSLIT", sub = "")
  # Step 3: Remove punctuation, digits, and all sapces
  names_column_new <- gsub("[[:punct:][:digit:]][[:space:]]", "", names_column_new)
  # Step 4: Create a new variable with only alphabetic characters
  names_column_new <- gsub("[^[:alpha:]]", "", names_column_new)
  return(names_column_new)
}
```


# Perform data cleaning on dfA using the clean_names function
```{r echo=TRUE, message=FALSE, warning=FALSE}
dfA <- dfA |>
 dplyr::mutate_at(dplyr::vars(FIRST_NAME, LAST_NAME), list(new = clean_names)) |>
 dplyr::mutate(FN = FIRST_NAME_new, LN = LAST_NAME_new)
```


# Perform data cleaning on dfB using the clean_names function
```{r echo=TRUE, message=FALSE, warning=FALSE}
dfB <- dfB |>
 dplyr::mutate_at(dplyr::vars(RecipientNameFirst, RecipientNameLast), list(new = clean_names)) |>
 dplyr::mutate(FN = RecipientNameFirst_new, LN = RecipientNameLast_new)


rm(clean_names)
```


# Create no name list
```{r echo=TRUE, message=FALSE, warning=FALSE}
NoNameList <- c(
  "NICKNAME",
  "NOFAMILYNAME",
  "NOFIRSTNAME",
  "NOLASTNAME",
  "NOMIDDLENAME",
  "NONAME",
  "NO",
  "UNKNOWN",
  "UNK",
  "UN",
  "NA"
)
# 
# # Blank out the names in the data if they match any of the strings in the NoNameList
dfA <- dfA |>
 dplyr::mutate(FN = dplyr::case_when(
                              FN %in% NoNameList~ "",
                              TRUE ~ FN),
               LN = dplyr::case_when(
                                LN %in% NoNameList ~ "",
                                TRUE ~ LN))

dfB <- dfB |>
 dplyr::mutate(FN = dplyr::case_when(
                              FN %in% NoNameList~ "",
                              TRUE ~ FN),
                LN = dplyr::case_when(
                                LN %in% NoNameList ~ "",
                                TRUE ~ LN))
rm(NoNameList)
```


# Delete rows that have missing First Name (FN), Last Name (LN), or Date of Birth (DOB).
```{r echo=TRUE, message=FALSE, warning=FALSE}
dfA <- dfA |>
  dplyr::filter(!is.na(FN) & FN != "" & !is.na( LN) & LN != "" & !is.na(DOB) )

dfB <- dfB |>
  dplyr::filter(!is.na(FN) & FN != "" & !is.na( LN) & LN != "" & !is.na(DOB) )
```


# Exact Matching 
```{r echo=TRUE, message=FALSE, warning=FALSE}
Exact.match <- merge(dfA, dfB, by=c("FN","LN","DOB", "gender"))
nrow(Exact.match)
```


# Using the fastLink R package for record linkage
```{r echo=TRUE, message=FALSE, warning=FALSE}
matches.out <- fastLink::fastLink(
                 dfA = dfA, dfB = dfB, 
                 # Specify the vector of variable names to be used for matching.
                 # These variable names should exist in both dfA and dfB
                 varnames = c('FN', 'LN', 'dob_day', 'dob_month', 'dob_year', 'gender'),
                 # Specify which variables among varnames should be compared using string distance
                 stringdist.match = c('FN', 'LN'),
                 # Specify which variables present in  stringdist.match can be partially matched
                 partial.match = c('FN', 'LN'),
                 # Specify which variables should be matched numerically
                 # Must be a subset of 'varnames' and must not be present in 'stringdist.match'.
                 numeric.match = c('dob_day', 'dob_month', 'dob_year'),
                 # Specify the number of CPU cores to utilize (parallel processing). The default value is NULL.
                 n.cores = numCores, 
                 return.all = TRUE,
                 return.df = TRUE
)
```


# Confusion Matrice
```{r echo=TRUE, message=FALSE, warning=FALSE}
# The output from fastLink() will be a list  that includes a matrix where each row
# is a match with the relevant indices of dfA (column 1) and dfB (column 2).

fastLink::confusion(matches.out, threshold = 0.98)

# Examine the EM object:
#matches.out$EM
```


# Summarize the accuracy of the match:
```{r echo=TRUE, message=FALSE, warning=FALSE}
# each column gives the match count, match rate,
# false discovery rate (FDR) and false negative rate (FNR)
# under different cutoffs for matches based on the posterior 
# probability of a match. 
summary(matches.out)
```


# Get fuzzy matches using the results from fastLink
```{r echo=TRUE, message=FALSE, warning=FALSE}
# A threshold of 0.98 is set for match classification
fuzzy_matches <-  fastLink::getMatches(dfA, dfB, 
                            fl.out = matches.out, 
                            threshold.match = 0.98)
# dput(names(fuzzy_matches))
# c("FIRST_NAME", "LAST_NAME", "BIRTH_DATE", "gender", "row_idA", 
# "dob_day", "dob_month", "dob_year", "DOB", "FIRST_NAME_new", 
# "LAST_NAME_new", "FN", "LN", "RecipientNameFirst", "RecipientNameLast", 
# "RecipientDateOfBirth", "row_idB", "RecipientNameFirst_new", 
# "RecipientNameLast_new", "gamma.1", "gamma.2", "gamma.3", "gamma.4", 
# "gamma.5", "gamma.6", "posterior")

## Display nicely...
fuzzy_matches |> 
  dplyr::mutate( posterior = round(posterior*100, digits = 3)) |>
  dplyr::select("posterior", "FIRST_NAME", "LAST_NAME","BIRTH_DATE", "RecipientNameFirst", "RecipientNameLast","RecipientDateOfBirth" ) |>
  knitr::kable()
```

# Blocking data
```{r echo=TRUE, message=FALSE, warning=FALSE}
blockdata_out <- fastLink::blockData(dfA, dfB, 
        varnames = c('FN', 'LN', 'dob_day', 'dob_month', 'dob_year'),
        # window blocking for numeric variables, where a given observation in dataset A
        #will be compared to all observations in dataset B where the value of the blocking
        #variable is within ±K of the value of the same variable in dataset A. 
        #The value of K is the size of the window
        # window.block = "birthyear", 
        # window.size = 1,
        #  using k-means clustering, so that similar values of string and numeric
        # variables are blocked together. 
        kmeans.block = "FN", 
        nclusters = 2)


```


# Aggregating Multiple Matches Together

```{r echo=TRUE, message=FALSE, warning=FALSE}
#Often, we run several different matches for a single data set - 
#for instance, when blocking by gender or by some other criterion to reduce the 
#number of pairwise comparisons. Here, we walk through how to aggregate those 
#multiple matches into a single summary. 
blockgender_out <- fastLink::blockData(dfA, dfB, varnames = "gender")
## Subset dfA into blocks
dfA_block1 <- dfA[blockgender_out$block.1$dfA.inds,]
dfA_block2 <- dfA[blockgender_out$block.2$dfA.inds,]

## Subset dfB into blocks
dfB_block1 <- dfB[blockgender_out$block.1$dfB.inds,]
dfB_block2 <- dfB[blockgender_out$block.2$dfB.inds,]

## Run fastLink on each
fl_out_block1 <- fastLink::fastLink(
  dfA_block1, dfB_block1,
  varnames = c('FN', 'LN', 'dob_day', 'dob_month', 'dob_year'), 
  n.cores = numCores
)
fl_out_block2 <- fastLink::fastLink(
  dfA_block2, dfB_block2,
  varnames = c('FN', 'LN', 'dob_day', 'dob_month', 'dob_year'), 
  n.cores = numCores
)

#Here, we run fastLink() on the subsets of data defined by blocking on gender in the previous section:
  
## Run fastLink on each
link.1 <- fastLink::fastLink(
    dfA_block1, dfB_block1,
    varnames = c('FN', 'LN', 'dob_day', 'dob_month', 'dob_year'), 
    n.cores = numCores
  )

link.2 <- fastLink::fastLink(
  dfA_block2, dfB_block2,
  varnames = c('FN', 'LN', 'dob_day', 'dob_month', 'dob_year'), 
  n.cores = numCores
)

#To aggregate the two matches into a single summary, we use the aggregateEM() function as follows:
agg.out <- fastLink::aggregateEM(em.list = list(link.1, link.2))

summary(agg.out )

# fuzzy_matches2 <-  fastLink::getMatches(dfA, dfB, 
#                             fl.out = agg.out, 
#                             threshold.match = 0.98)
# dput(names(fuzzy_matches2))
# ## Display nicely...
# fuzzy_matches2 |> 
#   dplyr::mutate( posterior = round(posterior*100, digits = 3)) |>
#   dplyr::select("posterior", "FIRST_NAME", "LAST_NAME","BIRTH_DATE", "RecipientNameFirst", "RecipientNameLast","RecipientDateOfBirth" ) |>
#   knitr::kable()


```




 
